---
title: "An Overview of the Tate Art Museum"
output: html_notebook
---

<style type = "text/css">
  body{
  font-family: math;
  font-size: 12pt;
  }
</style>


## Introduction

The Tate Art Museum, located in London, is home to the national collection of 16th century of British art and is its national modern art collection, being one of the largest museums of contemporary and modern art in the world. Yet, just what are the specialties of their art exhibits? 
  
  * What eras of art do they have? 
  * What is their largest piece? Their smallest?
  * How many artists haven't passed yet?
  * Where even are the exhibited artists from?


## Loading the Tidyverse and CSV Files

As a preface, both read_csv files are needed. The _artwork.csv_ includes the specifics of the artwork, and the _artist.csv_ the details of the artist.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
artwork <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-12/artwork.csv')
artists <- readr::read_csv("https://github.com/tategallery/collection/raw/master/artist_data.csv")
```

```{r}
artwork
```

```{r}
artists
```