---
title: "_An Overview of the Tate Art Museum_"
author: Jena Baxter - "jbaxter1244@floridapoly.edu"
output: html_notebook
---

<style type = "text/css">
  body{
  font-family: math;
  font-size: 13pt;
  }
</style>


# Introduction

The Tate Art Museum, located in London, is home to the national collection of 16th century of British art and is its national modern art collection, being one of the largest museums of contemporary and modern art in the country. Yet, just what are the specialties of their art exhibits? 
  
  * How varied are the mediums of the pieces?
  * What is their largest piece?
  * Where even are the exhibited artists from?

With this project, it aims to observe a new look at an art museum and delve closer to the more minute pieces of the artwork.

# Loading the Tidyverse and CSV Files

As a preface, both read_csv files are needed. The _artwork.csv_ includes the specifics of the artwork, and the _artist.csv_ the details of the artist.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
artwork <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-12/artwork.csv')
artists <- readr::read_csv("https://github.com/tategallery/collection/raw/master/artist_data.csv")
```

```{r}
artwork
```

```{r}
artists
```



## Era Organization

As art pieces are denoted typically by their stylistic choices, sorting the pieces by art era allows any reader to easily see what pieces originated during what time. While some of the years of creation are unknown and labeled as such, the rest of the vector gives a broad look at the showcased time periods. Looking through the _artwork_ set shows that there's loads of unknown years for some pieces! However, looking through the _mediums_ and at the actual _urls_ of the pieces themselves on the _artwork_ dataset shows that many may have simply been sketches or smaller works. Thus, they can't be as accurately dated as pieces that were meant to be seen by the public.

These classifications also may be a mix of other eras as it was simply easier to classify some in groups rather than alone. In order, the eras would be _Mannerism_, _Baroque_, _Rococo_, _Classicism_, _Romanticism_, _Realism_, _Impressionism_, _Symbolism_, _Art Nouveau_, _Expressionism_, _Surrealism_ (which is an overarching name to _Cubism_, _Futurism_, and _Dadaism_), _Pop Art_, and finally our current art era, _Modernism_.

```{r}
#function that sorts artists by era
art_eras <- artwork %>%
  mutate(
    era = case_when(
          year %in% c(1520:1599) ~ "mannerism",
          year %in% c(1600:1780) ~ "baroque_rococo",
          year %in% c(1781:1849) ~ "class_romance_real",
          year %in% c(1850:1889) ~ "impressionism",
          year %in% c(1890:1905) ~ "symbol_nouveau_express",
          year %in% c(1906:1924) ~ "surrealism",
          year %in% c(1925:1954) ~ "abstract",
          year %in% c(1955:1969) ~ "popart",
          year %in% c(1970:2020) ~ "modern",
          TRUE ~ "unknown"
    ) ) 
art_eras  
```


For example, what if you wanted to see the art made during the _surrealism_ era?

```{r}
filter(art_eras, era == "surrealism")
```

This prompts a list of artists and their respective work from the years 1906 to 1924 with the "era" category of surrealism from the _artwork.csv_. 

Doing this with the _artists.csv_ would produce a similar result, but is ultimately made moot due to the name of the artists already being displayed.





# Prior Exploration

Before answering any of the proposed questions, a brief overview of both datasets is necessary:

```{r}
summary(artists)
```

```{r}
summary(artwork)
```

## Data Exploration

As both datasets are incredibly large, organizing it by a much smaller margin makes the data easier to handle. Broader visualizations allow a simple overview of what we're handling.


### Univariate Analysis

Below is a few basic visuals of the datasets. Specializing and focusing on the broader variables will make combining them for more complex answers later on easier.

#### Art Periods

Through a brief overview of one of the many variables, we can check to see when exactly most pieces come from - era-wise.

```{r}
filter(count(art_eras, era)) %>% 
  ggplot() +
  geom_bar(aes(x = n, y = era, color = era), stat = 'identity')
```
The bar graph above shows that over 40,000 of our pieces were made during _Classicism_, _Romanticism_, or _Realism_! Being as the museum is in Europe, it's understandable that most of their pieces come from art periods of high prominence in an older Europe.



#### Genders

As another variable includes gender, I expect to see a prominence of male artists to female due to the age of most of the art as well.

```{r}
filter(count(artists, gender)) %>% 
  ggplot() +
  geom_histogram(aes(x = n, y = gender, color = gender), stat = 'identity')
```
And the assumption is entirely correct. Just under 3,000 artists are male, while just over 500 are female. However, we also are left with a very small amount of unknown artists more than likely due to a lack of updated data on the set's end or a genuine lack of knowledge for the artist.



### Bivariate Analysis

Since many variables in this set include long names, I've opted to leave those variables for other visualizations and instead pose a different look at a combination of data

#### Deaths

We can also see how many artists have died in what year with their genders. With the newfound knowledge that most of our pieces are a few decades or centuries old we can assume that most have passed by now and will be male.

```{r}
artists %>% 
  ggplot() +
  geom_point(aes(x = gender, y = yearOfDeath, color = gender, alpha = 0.25))
```

And the assumption is correct. The majority of artists have passed and were almost always male, the removed values the warning mentions referring to those who are living.


And seen below are artists who've been the more recent of the deaths. We can also filter those recent deaths by gender as well and see our already small amount of female artists is even shorter. I can imagine if the list were of those who were living, it'd be minuscule.

```{r}
filter(artists, yearOfDeath > 2000) %>% 
  filter(gender == "Female")
```






# Data Analysis



### How varied are the mediums of the pieces?


A commonality between pieces is usually a medium - what they're actually made from. The easiest way to see the _amount_ of each medium is by organizing them by such and having a count of each one. By placing them in a vector, we can also pull out the least and greatest used mediums.

```{r}
mediums <- artwork %>% 
  group_by(medium) %>%
  count(medium)
mediums
```

We can see here this list is incredibly extensive! The data shows that the museum went to extensive lengths and care to specify everything in incredible detail, as almost everything in the set is unique. Note as well that some pieces might be made of _multiple pages_, hence why the first few pages of the set include numbers.

For example, [in this image](http://www.tate.org.uk/art/images/work/T/T07/T07711_8.jpg), titled "Self-Portrait Made Using a House Converted into a Pinhole Camera", is actually ten different photographs placed together to make one whole piece. 




### What is their largest piece?

Being as varied as the art world, organizing their pieces via size is a metric that gives someone who can't be there in person a better understanding of its size.

```{r}
art_eras %>% 
  arrange(desc(dimensions)) %>% 
  select(title, dimensions, medium, artist)
```

From this sorted table we can see our top piece doesn't even have a set size - but going to the _medium_ column tells us it's actually two video projections! This means, that since the piece _"Consolation Service"_ can be varied in size, the second and actual largest piece would be _"Visceral Canker"_.

It's also noteworthy to add that some pieces later on aren't exactly in order due to the sizes including depth as well. For example, [this piece](https://www.tate.org.uk/art/images/work/T/T07/T07056_10.jpg) called _"Aurocthonous AAAAHHHH"_ is a projection onto colored wood and fabric. Its dimensions _include the projector as well_, thus making the piece wider than it actually is.




## Where even are the exhibited artists from?

With all the vast mediums and countless pieces, where did these creators even start out? 

```{r}
birthplace <- artists %>% 
  group_by(placeOfBirth) %>% 
  count(placeOfBirth) 
birthplace
```


The list above sorts by its specified _state and country_. We can see that many _countries_ are repeated, but there's less states repeated overall. Regardless, the list shows artists came from places as far as Brazil/Brasil and Cuba. Resorting them below gives a greatest-to-least overview of where most were born.


```{r}
birthplace %>% 
  arrange(desc(n))
```

Unsurprisingly, most are born in London specifically - it's where the museum _is_ located after all. However, there's a good handful of artists without a birthplace listed. In cases like this, it's more than likely the museum simply didn't catalog or know their birthplace much like the year that some pieces were made.

The importance of asking their locations is more than expected. Knowing where an artist came from gives insight of what kind of culture influenced their work into what it became. For example, though most _modern_ art is avant-garde and confusing, there's always some message behind it that the creator may have came up with, or maybe certain materials and colors were used to reference something from their home country.





# Conclusion

Through all of this analyzing, I've felt like the only thing missing was an actual trip to the pieces themselves. We've learned that the majority of the art made was _Classicism_, _Romantic_, or, _Real_, most of the artists were male with less than five hundred women artists total, that most of these artists have since passed, many of the mediums used are incredibly unique and specialized, the largest piece of artwork can be _varied in size_, and that most artists did originate from London itself. Even with all that, however, I had questions that were much smaller answered for me even if it wasn't intentional. I was able to apply the usage of _"mutate"_ and _"case_when"_ to actually make the expansive list that helped create the actual eras of art these pieces came from . I feel my report covers not only the grandiose overview of the Tate Art Museum, but was an incredibly enjoyable display of the work and tools I learned over this summer - both in and out of class.


# Future Ideas

As an art museum normally would, they update their records and exhibits when new material arrives. If the NAs of either fill in, or new exhibits are displayed and added to the dataset, I'd happily look into these sets again. Other questions or ideas I had were organizing by year of birth and charting it so, but at the current moment for brevity I left it with death years and simply seeing where they were born. If I also had more time, I'd make a function that changes the metric of the sizes for a bit more ease of understanding just how big the pieces are.

While there's a significant amount more I could have talked on, I feel going just in depth enough to be left with less questions than I began is a wonderful way to end this exploration. More information of the Tate Art Museum sets can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-12/readme.md) and the actual Tate Art Museum website can be found [here](https://www.tate.org.uk/).


# Citations

The datasets used can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-12/readme.md) and also above, and my reference for all the years used to make the dataset _"art_eras"_ can be found [here](https://artincontext.org/art-periods/) with explanations of every era and those uncluded.















